Detailed disease progression of 88 patients hospitalized with Covid-19 in the Czech Republic: A dataset and exploratory analysis

### 2020-09-12 manuscript/hmm.Rmd
More HMM models, work on manuscript

```{r}
series_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  rename(.time = day, .serie = patient_id, .observed = breathing) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup()
```

@@

```{r}
hcq_m5_formula <-  ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id) + mo(comorbidities_sum) * rate_group
hcq_m5 <- brmshmmdata(hcq_m5_formula, series_data, ☻
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

>>>>

```{r}
series_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup()

```

@@

```{r}
hcq_m5_formula <-  ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id) + mo(comorbidities_sum) * rate_group
hcq_m5 <- brmshmmdata(hcq_m5_formula, series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

@@

## hcq_m6 - D-dimer

```{r}
hcq_m6_series_data <- series_data %>% 
  group_by(.serie) %>%
  filter(any(!is.na(d_dimer))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "d_dimer", new_column_name = "peak_d_dimer", initial_value = 0) %>%
  mutate(log_peak_d_dimer = log(peak_d_dimer + 1) - 4)

hcq_m6_initial_states <- hcq_m6_series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

hcq_m6_formula <-  ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id) + log_peak_d_dimer * rate_group

hcq_m6 <- brmshmmdata(hcq_m6_formula, hcq_m6_series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m6_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m6_fit <- brmhmm(hcq_m6, cache_file = paste0(cache_dir,"/hcq_m6.rds"))

summary(hcq_m6_fit$brmsfit)
```

```{r}
hcq_m6_epred_rect <- posterior_epred_rect(hcq_m6_fit)
hcq_m6_predicted_rect <- posterior_epred_to_predicted(hcq_m6_fit, hcq_m6_epred_rect)
hcq_m6_predicted <- posterior_rect_to_long(hcq_m6_fit, hcq_m6_predicted_rect)
```

```{r}
hcq_m6_predicted_df <- posterior_long_to_df(hcq_m6_fit$data, hcq_m6_predicted)

all_ids <- hcq_m6_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m6_predicted_df, all_ids[series_id], hcq_m6_fit$data) %>% print()

```{r}
pp_check_last_state(hcq_m6_fit, hcq_m6_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m6_fit, hcq_m6_predicted_rect)
pp_check_transitions_direction(hcq_m6_fit, hcq_m6_predicted_rect, states_from = 2:6)
```

## hcq_m7 - IL-6

```{r}
hcq_m7_series_data <- series_data %>% 
  group_by(.serie) %>%
  filter(any(!is.na(IL_6))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "IL_6", new_column_name = "peak_IL_6", initial_value = 1) %>%
  mutate(log_peak_IL_6 = log(peak_IL_6))

hcq_m7_initial_states <- hcq_m7_series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

hcq_m7_formula <-  ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id) + log_peak_IL_6 * rate_group

hcq_m7 <- brmshmmdata(hcq_m7_formula, hcq_m7_series_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m7_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m7_fit <- brmhmm(hcq_m7, cache_file = paste0(cache_dir,"/hcq_m7.rds"))

summary(hcq_m7_fit$brmsfit)
```

```{r}
hcq_m7_epred_rect <- posterior_epred_rect(hcq_m7_fit)
hcq_m7_predicted_rect <- posterior_epred_to_predicted(hcq_m7_fit, hcq_m7_epred_rect)
hcq_m7_predicted <- posterior_rect_to_long(hcq_m7_fit, hcq_m7_predicted_rect)
```

```{r}
hcq_m7_predicted_df <- posterior_long_to_df(hcq_m7_fit$data, hcq_m7_predicted)

all_ids <- hcq_m7_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m7_predicted_df, all_ids[series_id], hcq_m7_fit$data) %>% print()

```{r}
pp_check_last_state(hcq_m7_fit, hcq_m7_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m7_fit, hcq_m7_predicted_rect)
pp_check_transitions_direction(hcq_m7_fit, hcq_m7_predicted_rect, states_from = 2:6)
```

### 2020-09-12 manuscript/manuscript.Rmd
More HMM models, work on manuscript

```{r setup}
devtools::load_all()
library(tidyverse)
theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()
wide <- prepare_marker_data_wide(data)
```

@@

>>>>

```{r setup, echo=FALSE, message=FALSE, warning = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)

devtools::load_all()
library(tidyverse)
theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()
wide <- prepare_marker_data_wide(data)
```

@@

```{r}

data_table_1 <- rbind(
  data$patient_data %>% mutate(group = "All"),
  data$patient_data %>% mutate(group = if_else(ever_hcq, "HCQ", "No HCQ")),
  data$patient_data %>% filter(any_d_dimer) %>% mutate(group = "Measured D-dimer"),
  data$patient_data %>% filter(any_IL_6) %>% mutate(group = "Measured IL-6")
  )

sum_percent <- function(x) {
  paste0(sum(x, na.rm = TRUE), " (", scales::percent(mean(x, na.rm = TRUE), accuracy = 1), ")")
}

mean_iqr <- function(x) {
  paste0(round(mean(x, na.rm = TRUE)), " (", round(quantile(x, 0.25, na.rm = TRUE)), " - ", round(quantile(x, 0.75, na.rm = TRUE)), ")")  
}

data_table_1 %>% group_by(group) %>% 
  summarise(
    N = n(),
    `Distinct sites` = length(unique(hospital_id)),
    Female = sum_percent(sex == "F"),
    `Age (mean, IQR)` = mean_iqr(age),
    `Admitted for Covid` = sum_percent(admitted_for_covid),
    `Took Azithromycin` = sum_percent(ever_az),
    `Ischemic Heart Disease` = sum_percent(ischemic_heart_disease),
    `Takes antihypertensives` = sum_percent(has_hypertension_drugs),	
    `Heart Failure` = sum_percent(heart_failure),
    COPD = sum_percent(COPD),
    Asthma = sum_percent(asthma),
    `Other lung disease` = sum_percent(other_lung_disease),
    Diabetes = sum_percent(diabetes),
    `Renal Disease` = sum_percent(renal_disease),
    `Liver Disease` = sum_percent(liver_disease),
    `Smoking` = sum_percent(smoking),
    `BMI (mean, IQR)` = mean_iqr(BMI),
    .groups = "drop"
  ) %>%
  mutate(group = factor(group, levels= c("All", "HCQ", "No HCQ", "Measured D-dimer","Measured IL-6"), ordered = TRUE)) %>%
  arrange(group) %>%
  as.data.frame() %>%
  column_to_rownames("group") %>%
  t() %>% 
  knitr::kable()
```
